---
title: "Take Home Exercise 02"
author: "Henry Low"
date: "Sep 25 2024"
date-modified: "last-modified"
execute:
  evalu: true
  echo: true
  message: false
  freeze: true
format: html
editor: visual
date-format: "DD MM YYYY"
---

# Background

Road traffic accidents cause 1.19 million deaths and 20 to 50 million injuries annually, with most fatalities occurring in low- and middle-income countries (World Health Organisation (WHO)). Road traffic injuries are the leading cause of death for individuals aged 5–29 and place a heavy economic burden on nations, costing up to 3% of GDP. Thailand has some of the deadliest roads globally, with about 20,000 fatalities annually. From 2014 to 2021, accidents rose, especially on national highways, with black spots frequently found on straight roads and intersections.

# Objectives

Focusing in the Bangkok Metropolitan Region (BMR), Spatial Point Patterns Analysis (SPPA) will be used to explore and identify factors influencing road traffic accidents. Such facts are mainly behavioral and environmental, with this exercise aiming to incorporate temporal factors as well. Specific objectives are as follows:\
- To visualize the spatio-temporal dynamics of road traffic accidents in BMR using appropriate statistical graphics and geovisualization methods.\
- To conduct detailed spatial analysis of road traffic accidents using appropriate Network Spatial Point Patterns Analysis methods.\
- To conduct detailed spatio-temporal analysis of road traffic accidents using appropriate Temporal Network Spatial Point Patterns Analysis methods.

# Study Area

The study area will be Thailand (sf_polygon features) at [province](https://en.wikipedia.org/wiki/Provinces_of_Thailand) level.

# Data Sources

*(saved under 'data' folder)*\
[Thailand Road Accident 2019-2022](https://www.kaggle.com/datasets/thaweewatboy/thailand-road-accident-2019-2022) from Kaggle\
[Thailand Roads (OpenStreetMap Export) from HDX](https://s3.dualstack.us-east-1.amazonaws.com/production-raw-data-api/ISO3/THA/roads/lines/hotosm_tha_roads_lines_shp.zip)\
[Thailand - Subnational Administrative Boundaries from HDX](https://data.humdata.org/dataset/cod-ab-tha?)...[Click here to download here](https://data.humdata.org/dataset/d24bdc45-eb4c-4e3d-8b16-44db02667c27/resource/d0c722ff-6939-4423-ac0d-6501830b1759/download/tha_adm_rtsd_itos_20210121_shp.zip)


# 1 Setting Up


## 1.1 Loading R Packages

```{r}
pacman::p_load(tidyverse, spatstat, sf, spNetwork, tmap, classInt, viridis)
```


## 1.2 Loading Datasets

```{r}
# Load Thailand Province from HDX
th_bound_l1 <- st_read(dsn = "data/tha_adm_rtsd_itos_20210121_shp/", layer = "tha_admbnda_adm1_rtsd_20220121")
glimpse(th_bound_l1)

# Visualize the boundaries
plot(st_geometry(th_bound_l1))
```

```{r}
df <- data.frame(Value = unlist(th_bound_l1$ADM1_EN))
write_csv(df, "data/provinces.csv")
```


