---
title: "Take Home Exercise 01"
author: "Henry Low"
date: "Sep 2 2024"
date-modified: "last-modified"
execute:
  evalu: true
  echo: true
  message: false
  freeze: true
format: html
editor: visual
---

# Background

Road traffic accidents cause 1.19 million deaths and 20 to 50 million injuries annually, with most fatalities occurring in low- and middle-income countries (World Health Organisation (WHO)). Road traffic injuries are the leading cause of death for individuals aged 5–29 and place a heavy economic burden on nations, costing up to 3% of GDP. Thailand has some of the deadliest roads globally, with about 20,000 fatalities annually. From 2014 to 2021, accidents rose, especially on national highways, with black spots frequently found on straight roads and intersections.

# Objectives

Focusing in the Bangkok Metropolitan Region (BMR), Spatial Point Patterns Analysis (SPPA) will be used to explore and identify factors influencing road traffic accidents. Such facts are mainly behavioral and environmental, with this exercise aiming to incorporate temporal factors as well. Specific objectives are as follows:\
- To visualize the spatio-temporal dynamics of road traffic accidents in BMR using appropriate statistical graphics and geovisualization methods.\
- To conduct detailed spatial analysis of road traffic accidents using appropriate Network Spatial Point Patterns Analysis methods.\
- To conduct detailed spatio-temporal analysis of road traffic accidents using appropriate Temporal Network Spatial Point Patterns Analysis methods.

# Data Sources

*(saved under 'data' folder)*\
[Thailand Road Accident 2019-2022](https://www.kaggle.com/datasets/thaweewatboy/thailand-road-accident-2019-2022) from Kaggle\
[Thailand Roads (OpenStreetMap Export) from HDX](https://s3.dualstack.us-east-1.amazonaws.com/production-raw-data-api/ISO3/THA/roads/lines/hotosm_tha_roads_lines_shp.zip)\
[Thailand - Subnational Administrative Boundaries from HDX](https://data.humdata.org/dataset/cod-ab-tha?)...[Click here to download here](https://data.humdata.org/dataset/d24bdc45-eb4c-4e3d-8b16-44db02667c27/resource/d0c722ff-6939-4423-ac0d-6501830b1759/download/tha_adm_rtsd_itos_20210121_shp.zip)

# 1. Setting Up

## 1.1 Loading R Packages

I will be using the following R packages:\
-`sf` package to perform geospatial wrangling tasks - `tidyverse` package for reading csv files, dataframe processing tasks - `tmap` for plotting tasks

```{r}
pacman::p_load(tidyverse, sf, tmap)
```

## 1.2 Loading Datasets

```{r}
# Load 2019-2020 Road Accidents Data
ra_tbl <- read_csv('data/thai_road_accident_2019_2022.csv')
# Check data
glimpse(ra_tbl)
```

```{r}
# Load 2019-2020 Road Accidents Data
ra_tbl <- read_csv('data/thai_road_accident_2019_2022.csv')
# Check data
glimpse(ra_tbl)
```

```{r}
# Load Thailand Roads from Openstreet Map
th_road <- st_read(dsn = "data/hotosm_tha_roads_lines_shp/", layer = "hotosm_tha_roads_lines_shp") 
```

```{r}
# Check data
# glimpse(th_road)
head(th_road, n=5)
```

```{r}
# Load Thailand Country from HDX
th_bound_l0 <- st_read(dsn = "data/tha_adm_rtsd_itos_20210121_shp/", layer = "tha_admbnda_adm0_rtsd_20220121")
glimpse(th_bound_l0)
plot(st_geometry(th_bound_l0))
```

```{r}
# Load Thailand Province from HDX
th_bound_l1 <- st_read(dsn = "data/tha_adm_rtsd_itos_20210121_shp/", layer = "tha_admbnda_adm1_rtsd_20220121")
glimpse(th_bound_l1)
plot(st_geometry(th_bound_l1))
```

```{r}
# Load Thailand Province from HDX
th_bound_l2 <- st_read(dsn = "data/tha_adm_rtsd_itos_20210121_shp/", layer = "tha_admbnda_adm2_rtsd_20220121") %>%
  st_transform(crs = 32647) %>%
  filter(ADM1_EN %in% c("Bangkok", "Nonthaburi", "Pathum Thani", 
                              "Samut Prakan", "Samut Sakhon", "Nakhon Pathom"))
glimpse(th_bound_l2)
plot(st_geometry(th_bound_l2))
```

```{r}
# Load Thailand Province from HDX
th_bound_l3 <- st_read(dsn = "data/tha_adm_rtsd_itos_20210121_shp/", layer = "tha_admbnda_adm3_rtsd_20220121") %>%
  st_transform(crs = 32647) %>%
  filter(ADM1_EN %in% c("Bangkok", "Nonthaburi", "Pathum Thani", 
                              "Samut Prakan", "Samut Sakhon", "Nakhon Pathom"))
glimpse(th_bound_l3)
plot(st_geometry(th_bound_l3))
```

```{r}
# Load Thailand Province from HDX
th_bound_lall <- st_read(dsn = "data/tha_adm_rtsd_itos_20210121_shp/", layer = "tha_admbndl_admALL_rtsd_itos_20220121") %>% 
  st_transform(crs = 32647) %>%
  filter(admLevel == 3)
glimpse(th_bound_lall)
plot(st_geometry(th_bound_lall))
```

```{r}
# Load Thailand Province from HDX
th_bound_lallndp <- st_read(dsn = "data/tha_adm_rtsd_itos_20210121_shp/", layer = "tha_admbndp_admALL_rtsd_itos_20220121") %>%
  st_transform(crs = 32647) %>%
  filter(ADM1_EN %in% c("Bangkok", "Nonthaburi", "Pathum Thani", 
                              "Samut Prakan", "Samut Sakhon", "Nakhon Pathom"))
glimpse(th_bound_lallndp)
plot(st_geometry(th_bound_lallndp))
```


```{r}
th_road_nonna <- th_road %>%
  filter(!is.na(highway), !is.na(name_en))
plot(st_geometry(th_road_nonna))
```

