---
title: "Hands-On Exercise 01"
author: "Henry Low"
date: "Aug 30 2024"
date-modified: "last-modified"
execute:
  evalu: true
  echo: true
  message: false
  freeze: true
format: html
editor: visual
---

# Chapter 1: Geospatial Data Science with R

## 1.1 Data Sources (saved under 'data' folder)

From [data.gov.sg](https://data.gov.sg/):\
- [Master Plan 2014 Subzone Boundary (Web) (SHP)](https://data.gov.sg/datasets?query=Master+Plan+2014+Subzone+Boundary&page=1&resultId=d_d14da225fccf921049ab64238ff473d9)\
- [Master Plan 2014 Subzone Boundary (Web) (KML)](https://data.gov.sg/datasets?query=Master+Plan+2014+Subzone+Boundary&page=1&resultId=d_5cb80a95445f236737f6bc2bfe5f159d)\
- [Pre-Schools Location (KML)](https://data.gov.sg/datasets?query=Pre-Schools+Location&page=1&resultId=d_a72bcd23e208d995f3bd4eececeaca43)

Others:\
[Cycling Path from LTADataMall](https://datamall.lta.gov.sg/content/dam/datamall/datasets/Geospatial/CyclingPath_Jul2024.zip)\
[Listings from Airbnb (listings.csv)](https://insideairbnb.com/get-the-data/)

## 1.2 Setting Up

### 1.2.1 Loading the R packages

```{r}
pacman::p_load(tidyverse, sf)
```

### 1.2.2 Importing Datasets

```{r}
# Load MP14_SUBSONZE_WEB_PL
mpsz = st_read(dsn = "data/MasterPlan2014SubzoneBoundaryWebSHP", layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
# Load Cycling Path
cyclingpath = st_read(dsn = "data/CyclingPath_Jul2024", layer = "CyclingPathGazette")
```

```{r}
# Load PreSchool 
preschool = st_read("data/PreSchoolsLocation.kml")
```

## 1.3 Data Exploration

## 1.4 Data Visualization

## 1.5 Projection

## 1.6 Transforming Aspatial Data

## 1.7 Geoprocessing 

# Chapter 2: Thematic Mapping and GeoVisualisation with R
